### 2018-11-14 ms/ms_prosop.tex
updated pdf, added some additional outputs to notebooks

The first case of prosopagnosia present from early childhood, now called developmental prosopagnosia (DP), was described by \citeauthor{McConachie_1976} in 1976. Unlike the prosopagnosia cases reported before \citep{Bodamer_1947, Benton_1968}, McConachie's patient showed no signs of brain damage that could explain the condition. It took 20 years until a second case of DP was reported \citep{Ariel_1996}, and hence the condition was considered to be very rare. However, with increased awareness of prosopagnosia in public, more people reporting symptoms of “face blindness” have come forward \citep{Duchaine_2006curr}. Today it is estimated that around 2\% of the general population presents with DP \citep{Kennerknecht_2008, Bowles_2009}. Similar to other people with lifelong prosopagnosia \citep{Sacks, Fine_2012}, JB reports that she was oblivious to the nature of her condition for the vast majority of her life. “It simply never occurred to me that one could recognize people only by their face,” JB told us. A few years ago she was watching a TV talk show, in which one of the guests was interviewed about his prosopagnosia. JB describes this moment as the one in which she immediately knew that she must have the same condition.

@@

DP is a condition of varying specificity and particularly the early cases were impaired in more general object recognition as well \citep{McConachie_1976, Ariel_1996}. Later, very “pure” cases, restricted to deficits in recognizing only the identity of faces were reported \citep{Nunn_2001, Bentin_1999, Duchaine_2005}. To exclude a more general form of agnosia, we performed a screening with living as well as non-living objects (cf. methods section), on which JB made no mistakes. In two fMRI tasks with faces, hands and landscapes as stimuli, JB showed prototypical activity in the lateral fusiform gyrus for faces (Fig.~\ref{fig:brain}b,c) and activity in medial parts of the ventral stream for landscapes (Fig.~\ref{fig:brain}b). These are the patterns which are predicted by the standard models of object processing \citep{Grill_Spector_2014} and indicate that JB's abilities to differentiate between faces and non-face objects are unimpaired. Such normal activation patterns in the fusiform gyrus for faces, lateral to the mid-fusiform sulcus \citep{Weiner_2014}, have been reported in previous cases with DP \citep{Hasson_2003, Avidan_2005, Furl_2011}. However, from discussions with JB we learned that faces as a superordinate category might pose a challenge for her when forming or accessing internal representations. For example, she often retains vivid recollections of her dreams, but they do not feature faces: “Whenever there is a person, instead of the face there is just a placeholder of some kind”. When reading fiction, JB claims that she always imagines the actions of the people described from a first-person perspective, which obviates imagining what the person performing that action might look like. Deficits in imagery for faces have also been found in some previous studies on DP \citep{Gr_ter_2009, Tree_2010}.
Furthermore, JB complained to us about problems with body perception, specifically correctly identifying her left and right hand. While JB is unequivocally right handed according to the Edinburgh handedness questionnaire \citep{Oldfield_1971}, she claims that she often uses the “wrong” (i.e. left) hand when initiating a movement and then has to correct herself and switch hands. In patients with damage to the fusiform gyrus, similar problems with body perception have been reported \citep{Grewe_2014}. Accordingly, JB's performance on the Bergen Right-Left Discrimination Test \citep{Ofte_2002} was two standard deviations (percentile 2; cf. Table~\ref{tab:summary}) below average, while her performance on a more general visuo-spatial task (subtests six to eight of the LPS-2; \citet{Horn}) was poor but still within the normal range (percentile 16). This might indicate that some body-specific visuo-spatial functions, which are neuroanatomically located close to the primary face processing areas, can also be impaired in certain cases of DP.

@@

JB reports that her condition is especially impairing when she has to learn new faces, as was the case when she held a job in sales and distribution: “If you have a job where you sit in your office and people come to you at previously appointed times, it is easy. But if you have to actively approach people, go to their offices, or make small talk in the hallway, you can't do it if you don't know who is who”. Her inability to recognize customers and colleagues therefore had many repercussions, even leading to the loss of some jobs.
To test JB's ability to learn and recognize new faces, we used the Cambridge Face Memory Test (CFMT; \citet{Duchaine_2006cfmt}), which is the most established test in DP research, and is considered most promising as a clinical diagnostic instrument \citep{Susilo_2013, Towler_2017}. It hides hair and other external features in the pictures and does not show faces simultaneously but consecutively, so feature matching is not as easily possible \citep{Duchaine_2006cfmt}. In the first part of the test, one face is shown for 2 seconds and after a brief pause 3 faces appear, the learned face being one of them. In this part, JB scored a perfect 100\% correct when trying to recognize the learned face (Fig.~\ref{fig:cfmt}). Afterwards, she explained to us that for each face she tried to find one or two characteristic features, verbalize them for herself and then search for them in subsequent pictures (e.g. “elongated chin”, “fair eyes”, “chubby cheeks”).

@@

While remembering faces seen only on one occasion is a difficult task per se, people with DP are also unable to recognize individuals they have known for years \citep{Ariel_1996, Nunn_2001, Jones_2001, Fine_2011}. Similarly, JB reported several occasions on which she passed her husband on the street without recognizing him. On another occasion, JB passed a large wall of mirrors in a department store and, for a brief moment, mistook her own reflection for somebody mimicking her movements. Also, JB noted that when she looks at pictures of herself, e.g. taken on a vacation, she is surprised every time she sees a picture of herself. “I cannot remember what I look like”, she told us.
Despite being unable to recognize her husband or herself, JB claims that there are a number of famous people she can recognize by their faces with high certainty. Studies in which participants with DP were asked to recognize famous faces have shown that they are significantly worse than controls in these tasks. However, their absolute scores are surprisingly high, as they usually can still identify around 30-40\% of the faces correctly \citep{Rivolta_2012, Garrido_2009}. One explanation is that those famous faces for which there has been more exposure might be easier to recognize \citep{Rivolta_2012}. However, this contradicts the observation that most cases of DP have difficulties in recognizing family members and themselves, despite lifelong exposure \citep{Sacks, Nunn_2001, Fine_2011, Duchaine_2011}.
To test JB's ability to recognize famous faces and directly address the role of exposure, we asked her to provide us with a list of people she thinks she can recognize from looking at their face alone. The list JB provided consisted of 14 celebrities, including Angela Merkel, Barack Obama, some (but notably not all) actors from the shows “How I Met Your Mother”, “House” and “Law and Order”, as well as some other politicians and actors. In our final experiment, there were 42 famous people (5 images per person), including the ones from JB's list, actors on the same shows who were not on the list, and other famous people (cf. Table~\ref{tab:famous}).

@@

All pictures were gray-scaled and prepared so that only inner features were visible. The task for JB was to first decide if she knew the person (familiarity) then if she could pick the context from which the person might be familiar and finally to write down the person's name (either the actual name or, in the case of actors, the name of the character they play).
JB's performance was below the normal range for familiarity, context and for naming (Fig.~\ref{fig:famous}), indicating a clear deficit in recognizing familiar faces. However, her performance was average or above-average when only considering the faces on the list she prepared. Regarding free naming, JB reached only 14\% correct for famous people not on her list, but performed 69\% correct for the famous faces who were on her list. This illustrates that JB has good introspection into her abilities: there are faces that she can recognize by their inner features alone with high reliability. For example, JB successfully recognized the characters Lily (5 out of 5), Marshall (4/5) and Barney (3/5) from the sitcom “How I Met Your Mother”, who were all previously on her list. The other two main protagonists Ted and Robin from the same show were not on her list and both were not recognized even once (both 0/5), although JB should have had roughly the same amount of exposure to them.
To get a better understanding of how JB performed that task, we later asked her to try and verbalize what makes recognizing certain faces easy. Regarding Ms. Merkel (4/5 correct namings), JB pointed out that Ms. Merkel seems to have a “characteristic facial expression that is kind of stiff”. Regarding Mr. Obama (4/5 correct namings), JB said that he is easy to recognize due to his “freckles”, a subtle feature which most observers (including the authors) probably do not notice but which prove useful for observers who heavily rely on recognizing individuals by single features.

@@

 \begin{figure*}[htbp]
 	\begin{minipage}{\textwidth}
 		\renewcommand{\familydefault}{\sfdefault}\normalfont
 		\centering
 		\includegraphics[width=\columnwidth]{../reports/figures/filterface_main_figure_mod.png}
 		\vspace*{-3mm}
 		\caption{Results for the filtered famous faces. Face stimuli were all shown in grey-scale and with peripheral features visible. The “original” condition consists of stimuli with no further manipulation; the high-pass filtered (HPF) version consists of faces which were filtered so that the edges of features are emphasized; the low-pass filtered (LPF) version consists of faces which were smoothed with a Gaussian kernel so that features are blurred and featural processing of faces is not possible. Each face has to be rated regarding familiarity, context and the person's name has to be given. Results of the control participants are shown in blue (con), JB's results for famous faces she did not preselect are shown in red and JB's results for faces which she previously listed as being recognizable are shown in green (JB*). Except for the familiarity ratings on LPF faces, where JB's performance is 2 SD below the controls (percentile 2), her performance on all tasks and for both preselected and not preselected faces was within the normal range. Picture of Barack Obama is in the public domain.}%
 		\label{fig:filtered}
 	\end{minipage}
 \end{figure*}

To better understand how JB uses the features of the face compared to their configuration (so-called holistic or configural processing; \citet{Maurer_2002}), we modified the famous faces task to show low-pass and high-pass filtered images. When applying a low-pass filter (Fig.~\ref{fig:filtered}), a single blurred feature, for example an eye, cannot be recognized as such in isolation, but only in the context of the whole face. Hence, using low-pass filtered images in recognition tasks is useful to test the integrity of configural face processing by preventing single-feature analyses. As we anticipated this task to be very difficult, we used the full images with peripheral information. In the task, JB was relatively unimpaired, scoring 48\% for naming of the blurred images and 84\% for the unaltered images. Therefore, JB's performance is surprisingly high, even when configural processing is allegedly the only viable strategy. From discussions with JB, we learned that she claims to be able to judge “the interplay of different face parts” when confronted with blurred images. While intact configural processing in DP has been reported before \citep{Duchaine_2000, Le_Grand_2006}, it raises the question of what mechanism can explain the deficits observed in DP and why the good performance even under adversarial conditions does not easily translate into everyday life. When asked about this, JB claimed that one difficulty in everyday life is that familiar faces will not pop out from a crowd. However, they often can be found with enough effort. For example, while JB might walk by her husband on the street, she might be able to find him at a designated spot. Similarly, JB might scan a lecture hall row by row and seat by seat to finally find a fellow student she is looking for. This “one face at a time” strategy might also allow for good performance in certain experimental settings, like the present task.

@@

Another aspect of face processing that is usually unimpaired in people with DP is the recognition of emotions, intentions or other social information, such as trustworthiness \citep{Todorov_2008, Duchaine_2009}. Despite her prosopagnosia, JB says she is very good at picking up subtle social cues, which also aid her in deciding whether someone might be familiar. “I can recognize if someone recognizes me; and then I act accordingly”. She also judges her skills to tell if her interaction partner is lying, agitated or uncomfortable as very high. “I think I have a very keen understanding of social signals. The way the eyes of a person change when they are uncomfortable, small pauses they make in their speech and certain subtle facial expressions”. In order to objectively assess her ability to derive social cues from faces, we used the “Reading the mind in the eyes” test, which measures the ability to imagine the mental states of others \citep{Baron_Cohen_2001}. In the test, only the eye region of a person's face is shown and the participant is asked to select one word out of four which best describes the mental state of the depicted person (for example, “ashamed”, “alarmed”, “bewildered”, “irritated”). On the test, JB scored 26 points out of 36 (72\% correct), which is at the lower end of the normal range (cf. Table~\ref{tab:summary}). The only other case of DP in the literature who performed the “Reading the mind in the eyes” test also showed poor performance, even significantly below the normal range \citep{Duchaine_2006cog}.
To further test JB's ability to identify facial expressions of emotion, we used a 7-way forced-choice basic expression recognition task, with happy, sad, angry, fearful, disgusted, surprised and neutral faces. Here, JB scored 88\% correct and was the second-best participant in our sample (Fig.~\ref{fig:emo}). This is in line with most studies indicating that individuals with DP are unimpaired in recognizing emotions in faces \citep{Duchaine_2003, Humphreys_2006}; but see \citep{Duchaine_2006cog} for deficits in emotion recognition).

@@

In addition to identifying intentions and emotions, JB is also convinced that she can reliably judge the attractiveness of faces, and a number of episodes from her life seem to corroborate this impression.
In one episode, JB recalled being at a discotheque with a female friend. She saw a man she found handsome, had a drink with him and they talked for some amount of time. The other day, she was at a cafe with the same friend, as a man walked by. She remarked to her friend that he also looks quite handsome, to which her friend replied that this was the same man she had drinks with the night before. The same thing happened a third time, only a few days afterwards, with the same man, in pretty much the same way. This provides anecdotal evidence that JB's judgments of attractiveness are consistent across time.
JB also reported that she likes “odd” faces and takes an immediate liking to anyone with some idiosyncratic features in their face. She finds people with green eyes or thick eyebrows immediately likable, predominantly because they are easier to identify and consequently interaction with them is considerably less stressful. “I like red hair a lot and take an immediate liking to anyone who has it,” JB told us. Similarly, other cases of DP described in the literature have remarked that they focus on the “worst features” of each acquaintance's face to remember them, keep company with “physically distinctive” people and claim that forgetting someone's face should be regarded a compliment, because good looks are not memorable \citep{Fine_2012}.
Increasing the averageness of a face normally increases its perceived attractiveness \citep{Little_2011}) while making it less distinctive. Accordingly, we prepared a task where faces with different degrees of averageness were presented (Fig.~\ref{fig:attract}). In each trial a pair of faces was shown, one more and one less strongly averaged, and JB had to make a choice which one she finds more attractive or more likable. Compared to a group of 46 female control participants, JB showed a prototypical preference for more average faces when it comes to attractiveness, but was the only person in the group to find less average faces more likable. The results show that in DP, ratings of attractiveness and likability might dissociate. The findings are both in line with studies showing normal attractiveness ratings for individuals with DP \citep{Carbon_2010}, as well as the until now incompatible observation that they report liking 'odd' faces more.

@@

	\renewcommand{\familydefault}{\sfdefault}\normalfont
	\centering
	\includegraphics[width=\columnwidth]{../reports/figures/overview_heatmap_mod.png}
	\vspace*{-3mm}
 		\caption{Comparison of JB's performance for all major tasks. Differences in z-scores for all pairings of tasks are reported, with positive scores indicating higher performance for the tasks listed in the rows, compared to the tasks listed in the columns. Differences larger than two standard deviations are highlighted by stronger colors and an annotation indicating the numerical difference in z-scores. The emerging pattern suggests that JB’s performance is impaired for tasks involving learning of multiple faces (CFMT 2 and 3), and the recognition of famous faces from their inner parts, compared to learning of a single face (CFMT 1) and the recognition of emotion, intention (Eyes test) and attractiveness in faces. CFMT: Cambridge Face Memory Test; Eyes test: Reading the mind in the eyes; BRLT: Bergen Right-Left Discrimination Test; LPS2-Visual: Leistungsprüfsystem 2 (“performance test system”, test of visuo-spatial skills).}%
 		\label{fig:heat}

@@

We have presented the case of JB, a 30-year old woman suffering from DP, a lifelong face recognition deficit. JB presents with a pure form of DP, showing strongly impaired recognition of identity for both unfamiliar and familiar faces, but no difficulties recognizing expressions, intentions or attractiveness from faces (cf. Table~\ref{tab:summary}, Fig.~\ref{fig:heat}). She can also recognize identity from uncropped images of faces with peripheral features, indicating preserved configural processing of faces with peripheral features.

@@

While DP certainly is a heterogeneous disorder \citep{Duchaine_2006curr, Le_Grand_2006} and it is difficult to generalize from the single case, the CFMT provided the most unequivocal results in the present study, as it allowed us to best delineate JB's performance from the control participants. The perfect performance on the first part of the CFMT, where only one face has to be learned and recognized, also suggests that JB’s difficulties cannot be due to problems with face perception. This part of the CFMT strongly dissociated from the other two parts, where multiple faces have to be learned and recognized (Figure 7). In line with the present results, the CFMT has been successfully used to identify DP in a number of previous studies (e.g. \citet{Duchaine_2006cfmt, Duchaine_2006cog}). JB also told us that she found the CFMT to reflect her problems most faithfully, because many faces need to be learned at once and even extreme conscientiousness did not allow her to perform well when using only single features. A test that has so far worked for the majority of patients in the literature is important, especially to improve the state of clinical diagnostics in DP \citep{Shah_2016, Bate_2017}.

@@

\subsection*{Cambridge Face Memory Test}
To evaluate the memory for unfamiliar faces, we used the CFMT. The test consists of three parts, (i) learning one face, (ii) learning six faces (iii) learning six faces degraded by noise \citep{Duchaine_2006cfmt}. The control sample consisted of 27 female participants (mean age 27, range 25-35) who were students at Bielefeld University and did not report any history of psychiatric or neurological disorders. These participants also served as controls for the famous face tasks, which is described below. The experiment was programmed and presented using PsychoPy2.

@@

\subsection*{Statistical Methods}
For each test, the percent of correct responses was computed as the ratio of JB’s score relative to the maximal achievable score. JB’s z-score for each test was computed based on the control sample’s distribution of scores, which was transformed to have zero mean and unit variance. The z-scores were converted into percentile scores using a cumulative distribution function. Inferential statistics were computed using the t-test method developed by Crawford \& Howell, as described in \citet{Crawford_2009}. Two-sided p-values were computed based on the t-value and the n-1 degrees of freedom (n being the size of the control sample). All computation were implemented using in-house software written in Python 2.7. A heatmap visualizing the differences of JB’s z-scores in all major tasks was created using Seaborn (seaborn.pydata.org).

>>>>

The first case of prosopagnosia present from early childhood, now called developmental prosopagnosia (DP), was described by \citeauthor{McConachie_1976} in 1976. Unlike the prosopagnosia cases reported before \citep{Bodamer_1947, Benton_1968}, McConachie's patient showed no signs of brain damage that could explain the condition. It took 20 years until a second case of DP was reported \citep{Ariel_1996}, and hence the condition was considered to be very rare. However, with increased awareness of prosopagnosia in public, more people reporting symptoms of “face blindness” have come forward \citep{Duchaine_2006curr}. Today it is estimated that around 2\% of the general population presents with DP \citep{Kennerknecht_2008, Bowles_2009}. Similar to other people with lifelong prosopagnosia \citep{Sacks, Fine_2012}, JB reports that she was oblivious to the nature of her condition for the vast majority of her life. \textit{“It simply never occurred to me that one could recognize people only by their face”}, JB told us. A few years ago she was watching a TV talk show, in which one of the guests was interviewed about his prosopagnosia. JB describes this moment as the one in which she immediately knew that she must have the same condition.

@@

DP is a condition of varying specificity and particularly the early cases were impaired in more general object recognition as well \citep{McConachie_1976, Ariel_1996}. Later, very “pure” cases, restricted to deficits in recognizing only the identity of faces were reported \citep{Nunn_2001, Bentin_1999, Duchaine_2005}. To exclude a more general form of agnosia, we performed a screening with living as well as non-living objects (cf. methods section), on which JB made no mistakes. In two fMRI tasks with faces, hands and landscapes as stimuli, JB showed prototypical activity in the lateral fusiform gyrus for faces (Fig.~\ref{fig:brain}b,c) and activity in medial parts of the ventral stream for landscapes (Fig.~\ref{fig:brain}b). These are the patterns which are predicted by the standard models of object processing \citep{Grill_Spector_2014} and indicate that JB's abilities to differentiate between faces and non-face objects are unimpaired. Such normal activation patterns in the fusiform gyrus for faces, lateral to the mid-fusiform sulcus \citep{Weiner_2014}, have been reported in previous cases with DP \citep{Hasson_2003, Avidan_2005, Furl_2011}. However, from discussions with JB we learned that faces as a superordinate category might pose a challenge for her when forming or accessing internal representations. For example, she often retains vivid recollections of her dreams, but they do not feature faces: \textit{“Whenever there is a person, instead of the face there is just a placeholder of some kind”}. When reading fiction, JB claims that she always imagines the actions of the people described from a first-person perspective, which obviates imagining what the person performing that action might look like. Deficits in imagery for faces have also been found in some previous studies on DP \citep{Gr_ter_2009, Tree_2010}.
Furthermore, JB complained to us about problems with body perception, specifically correctly identifying her left and right hand. While JB is unequivocally right handed according to the Edinburgh handedness questionnaire \citep{Oldfield_1971}, she claims that she often uses the \textit{“wrong”} (i.e. left) hand when initiating a movement and then has to correct herself and switch hands. In patients with damage to the fusiform gyrus, similar problems with body perception have been reported \citep{Grewe_2014}. Accordingly, JB's performance on the Bergen Right-Left Discrimination Test \citep{Ofte_2002} was two standard deviations below average (cf. Table~\ref{tab:summary}), while her performance on a more general visuo-spatial task (subtests six to eight of the LPS-2; \citet{Horn}) was poor but still within the normal range (percentile 16). This might indicate that some body-specific visuo-spatial functions, which are neuroanatomically located close to the primary face processing areas, can also be impaired in certain cases of DP.

@@

JB reports that her condition is especially impairing when she has to learn new faces, as was the case when she held a job in sales and distribution: \textit{“If you have a job where you sit in your office and people come to you at previously appointed times, it is easy. But if you have to actively approach people, go to their offices, or make small talk in the hallway, you can't do it if you don't know who is who”}. Her inability to recognize customers and colleagues therefore had many repercussions, even leading to the loss of some jobs.
To test JB's ability to learn and recognize new faces, we used the Cambridge Face Memory Test (CFMT; \citet{Duchaine_2006cfmt}), which is the most established test in DP research, and is considered most promising as a clinical diagnostic instrument \citep{Susilo_2013, Towler_2017}. It hides hair and other external features in the pictures and does not show faces simultaneously but consecutively, so feature matching is not as easily possible \citep{Duchaine_2006cfmt}. In the first part of the test, one face is shown for 2 seconds and after a brief pause 3 faces appear, the learned face being one of them. In this part, JB scored a perfect 100\% correct when trying to recognize the learned face (Fig.~\ref{fig:cfmt}). Afterwards, she explained to us that for each face she tried to find one or two characteristic features, verbalize them for herself and then search for them in subsequent pictures (e.g. \textit{“elongated chin”}, \textit{“fair eyes”}, \textit{“chubby cheeks”}).

@@

While remembering faces seen only on one occasion is a difficult task per se, people with DP are also unable to recognize individuals they have known for years \citep{Ariel_1996, Nunn_2001, Jones_2001, Fine_2011}. Similarly, JB reported several occasions on which she passed her husband on the street without recognizing him. On another occasion, JB passed a large wall of mirrors in a department store and, for a brief moment, mistook her own reflection for somebody mimicking her movements. Also, JB noted that when she looks at pictures of herself, e.g. taken on a vacation, she is surprised every time she sees a picture of herself. \textit{“I cannot remember what I look like”}, she told us.
Despite being unable to recognize her husband or herself, JB claims that there are a number of famous people she can recognize by their faces with high certainty. Studies in which participants with DP were asked to recognize famous faces have shown that they are significantly worse than controls in these tasks. However, their absolute scores are surprisingly high, as they usually can still identify around 30-40\% of the faces correctly \citep{Rivolta_2012, Garrido_2009}. One explanation is that those famous faces for which there has been more exposure might be easier to recognize \citep{Rivolta_2012}. However, this contradicts the observation that most cases of DP have difficulties in recognizing family members and themselves, despite lifelong exposure \citep{Sacks, Nunn_2001, Fine_2011, Duchaine_2011}.
To test JB's ability to recognize famous faces and directly address the role of exposure, we asked her to provide us with a list of people she thinks she can recognize from looking at their face alone. The list JB provided consisted of 14 celebrities, including Angela Merkel, Barack Obama, some (but notably not all) actors from the shows “How I Met Your Mother”, “House” and “Law and Order”, as well as some other politicians and actors. In our final experiment, there were 42 famous people (5 images per person), including the ones from JB's list, actors on the same shows who were not on the list, and other famous people (cf. Table~\ref{tab:famous}).

@@

All pictures were gray-scaled and prepared so that only inner features were visible. The task for JB was to first decide if she knew the person (familiarity) then if she could pick the context from which the person might be familiar and finally to write down the person's name (either the actual name or, in the case of actors, the name of the character they play).
JB's performance was below the normal range for familiarity, context and for naming (Fig.~\ref{fig:famous}, Table~\ref{tab:summary}), indicating a clear deficit in recognizing familiar faces. However, her performance was average or above-average when only considering the faces on the list she prepared. Regarding free naming, JB reached only 14\% correct for famous people not on her list, but performed 69\% correct for the famous faces who were on her list. This illustrates that JB has good introspection into her abilities: there are faces that she can recognize by their inner features alone with high reliability. For example, JB successfully recognized the characters Lily (5 out of 5), Marshall (4/5) and Barney (3/5) from the sitcom “How I Met Your Mother”, who were all previously on her list. The other two main protagonists Ted and Robin from the same show were not on her list and both were not recognized even once (both 0/5), although JB should have had roughly the same amount of exposure to them.
To get a better understanding of how JB performed that task, we later asked her to try and verbalize what makes recognizing certain faces easy. Regarding Ms. Merkel (4/5 correct namings), JB pointed out that Ms. Merkel seems to have a \textit{“characteristic facial expression that is kind of stiff”}. Regarding Mr. Obama (4/5 correct namings), JB said that he is easy to recognize due to his \textit{“freckles”}, a subtle feature which most observers (including the authors) probably do not notice but which prove useful for observers who heavily rely on recognizing individuals by single features.

@@

 \begin{figure*}[htbp]
 	\begin{minipage}{\textwidth}
 		\renewcommand{\familydefault}{\sfdefault}\normalfont
 		\centering
 		\includegraphics[width=\columnwidth]{../reports/figures/filterface_main_figure_mod.png}
 		\vspace*{-3mm}
 		\caption{Results for the filtered famous faces. Face stimuli were all shown in grey-scale and with peripheral features visible. The “original” condition consists of stimuli with no further manipulation; the high-pass filtered (HPF) version consists of faces which were filtered so that the edges of features are emphasized; the low-pass filtered (LPF) version consists of faces which were smoothed with a Gaussian kernel so that features are blurred and featural processing of faces is not possible. Each face has to be rated regarding familiarity, context and the person's name has to be given. Results of the control participants are shown in blue (con), JB's results for famous faces she did not preselect are shown in red and JB's results for faces which she previously listed as being recognizable are shown in green (JB*). Except for the familiarity ratings on HPF and LPF faces, where JB's performance is 2 SD below the controls, her performance on all tasks and for both preselected and not preselected faces was within the normal range. Picture of Barack Obama is in the public domain.}%
 		\label{fig:filtered}
 	\end{minipage}
 \end{figure*}

To better understand how JB uses the features of the face compared to their configuration (so-called holistic or configural processing; \citet{Maurer_2002}), we modified the famous faces task to show low-pass and high-pass filtered images. When applying a low-pass filter (Fig.~\ref{fig:filtered}), a single blurred feature, for example an eye, cannot be recognized as such in isolation, but only in the context of the whole face. Hence, using low-pass filtered images in recognition tasks is useful to test the integrity of configural face processing by preventing single-feature analyses. As we anticipated this task to be very difficult, we used the full images with peripheral information. In the task, JB was relatively unimpaired, scoring 48\% for naming of the blurred images and 85\% for the unaltered images. Therefore, JB's performance is surprisingly high, even when configural processing is allegedly the only viable strategy. From discussions with JB, we learned that she claims to be able to judge \textit{“the interplay of different face parts”} when confronted with blurred images. While intact configural processing in DP has been reported before \citep{Duchaine_2000, Le_Grand_2006}, it raises the question of what mechanism can explain the deficits observed in DP and why the good performance even under adversarial conditions does not easily translate into everyday life. When asked about this, JB claimed that one difficulty in everyday life is that familiar faces will not pop out from a crowd. However, they often can be found with enough effort. For example, while JB might walk by her husband on the street, she might be able to find him at a designated spot. Similarly, JB might scan a lecture hall row by row and seat by seat to finally find a fellow student she is looking for. This \textit{“one face at a time”} strategy might also allow for good performance in certain experimental settings, like the present task.

@@

Another aspect of face processing that is usually unimpaired in people with DP is the recognition of emotions, intentions or other social information, such as trustworthiness \citep{Todorov_2008, Duchaine_2009}. Despite her prosopagnosia, JB says she is very good at picking up subtle social cues, which also aid her in deciding whether someone might be familiar. \textit{“I can recognize if someone recognizes me; and then I act accordingly”}. She also judges her skills to tell if her interaction partner is lying, agitated or uncomfortable as very high. \textit{“I think I have a very keen understanding of social signals. The way the eyes of a person change when they are uncomfortable, small pauses they make in their speech and certain subtle facial expressions”}. In order to objectively assess her ability to derive social cues from faces, we used the “Reading the mind in the eyes” test, which measures the ability to imagine the mental states of others \citep{Baron_Cohen_2001}. In the test, only the eye region of a person's face is shown and the participant is asked to select one word out of four which best describes the mental state of the depicted person (for example, “ashamed”, “alarmed”, “bewildered”, “irritated”). On the test, JB scored 26 points out of 36 (72\% correct), which is at the lower end of the normal range (cf. Table~\ref{tab:summary}). The only other case of DP in the literature who performed the “Reading the mind in the eyes” test also showed poor performance, even significantly below the normal range \citep{Duchaine_2006cog}.
To further test JB's ability to identify facial expressions of emotion, we used a 7-way forced-choice basic expression recognition task, with happy, sad, angry, fearful, disgusted, surprised and neutral faces. Here, JB scored 88\% correct and was the second-best participant in our sample (Fig.~\ref{fig:emo}). This is in line with most studies indicating that individuals with DP are unimpaired in recognizing emotions in faces \citep{Duchaine_2003, Humphreys_2006}; but see \citep{Duchaine_2006cog} for deficits in emotion recognition).

@@

In addition to identifying intentions and emotions, JB is also convinced that she can reliably judge the attractiveness of faces, and a number of episodes from her life seem to corroborate this impression.
In one episode, JB recalled being at a discotheque with a female friend. She saw a man she found handsome, had a drink with him and they talked for some amount of time. The other day, she was at a cafe with the same friend, as a man walked by. She remarked to her friend that he also looks quite handsome, to which her friend replied that this was the same man she had drinks with the night before. The same thing happened a third time, only a few days afterwards, with the same man, in pretty much the same way. This provides anecdotal evidence that JB's judgments of attractiveness are consistent across time.
JB also reported that she likes \textit{“odd”} faces and takes an immediate liking to anyone with some idiosyncratic features in their face. She finds people with green eyes or thick eyebrows immediately likable, predominantly because they are easier to identify and consequently interaction with them is considerably less stressful. \textit{“I like red hair a lot and take an immediate liking to anyone who has it”}, JB told us. Similarly, other cases of DP described in the literature have remarked that they focus on the “worst features” of each acquaintance's face to remember them, keep company with “physically distinctive” people and claim that forgetting someone's face should be regarded a compliment, because good looks are not memorable \citep{Fine_2012}.
Increasing the averageness of a face normally increases its perceived attractiveness \citep{Little_2011}) while making it less distinctive. Accordingly, we prepared a task where faces with different degrees of averageness were presented (Fig.~\ref{fig:attract}). In each trial a pair of faces was shown, one more and one less strongly averaged, and JB had to make a choice which one she finds more attractive or more likable. Compared to a group of 46 female control participants, JB showed a prototypical preference for more average faces when it comes to attractiveness, but was the only person in the group to find less average faces more likable. The results show that in DP, ratings of attractiveness and likability might dissociate (Fig.~\ref{fig:attract}, Fig.~\ref{fig:heat}). The findings are both in line with studies showing normal attractiveness ratings for individuals with DP \citep{Carbon_2010}, as well as the until now incompatible observation that they report liking 'odd' faces more.

@@

	\renewcommand{\familydefault}{\sfdefault}\normalfont
	\centering
	\includegraphics[width=\columnwidth]{../reports/figures/overview_heatmap_mod.png}
	\vspace*{-3mm}
 		\caption{Comparison of JB's performance for all major tasks. Differences in z-scores for all pairings of tasks are reported, with positive scores indicating higher performance for the tasks listed in the rows, compared to the tasks listed in the columns. Differences larger than two standard deviations are highlighted by stronger colors and an annotation indicating the numerical difference in z-scores. The emerging pattern suggests that JB’s performance is impaired for tasks involving learning of multiple faces (CFMT 2 and 3), and the recognition of famous faces from their inner parts, compared to learning of a single face (CFMT 1) and the recognition of emotion, intention (Eyes test) and attractiveness in faces. CFMT: Cambridge Face Memory Test; Eyes test: Reading the mind in the eyes; BRLT: Bergen Right-Left Discrimination Test; LPS2-Visual: Leistungsprüfsystem 2 (“performance test system”), test of visuo-spatial skills.}%
 		\label{fig:heat}

@@

We have presented the case of JB, a 30-year old woman suffering from DP, a lifelong face recognition deficit. JB presents with a pure form of DP, showing strongly impaired recognition of identity for both unfamiliar and familiar faces, but no difficulties recognizing expressions, intentions or attractiveness from faces (cf. Table~\ref{tab:summary}, Fig.~\ref{fig:heat}). She can also recognize identity from uncropped images of faces, indicating preserved configural processing of faces with peripheral features.

@@

While DP certainly is a heterogeneous disorder \citep{Duchaine_2006curr, Le_Grand_2006} and it is difficult to generalize from the single case, the CFMT provided the most unequivocal results in the present study, as it allowed us to best delineate JB's performance from the control participants. The perfect performance on the first part of the CFMT, where only one face has to be learned and recognized, also suggests that JB’s difficulties cannot be due to problems with face perception. This part of the CFMT strongly dissociated from the other two parts, where multiple faces have to be learned and recognized (Fig.~\ref{fig:heat}). In line with the present results, the CFMT has been successfully used to identify DP in a number of previous studies (e.g. \citet{Duchaine_2006cfmt, Duchaine_2006cog}). JB also told us that she found the CFMT to reflect her problems most faithfully, because many faces need to be learned at once and even extreme conscientiousness did not allow her to perform well when using only single features. A test that has so far worked for the majority of patients in the literature is important, especially to improve the state of clinical diagnostics in DP \citep{Shah_2016, Bate_2017}.

@@

\subsection*{Cambridge Face Memory Test}
To evaluate the memory for unfamiliar faces, we used the CFMT. The test consists of three parts, (i) learning one face, (ii) learning six faces (iii) learning six faces degraded by noise \citep{Duchaine_2006cfmt}. The control sample consisted of 27 female participants (mean age 27, range 25-35) who were students at Bielefeld University and did not report any history of psychiatric or neurological disorders. These participants also served as controls for the famous face tasks, which are described below. The experiment was programmed and presented using PsychoPy2.

@@

\subsection*{Statistical Methods}
For each test, the percent of correct responses was computed as the ratio of JB’s score relative to the maximal achievable score. JB’s z-score for each test was computed based on the control sample’s distribution of scores, which was transformed to have zero mean and unit variance. The z-scores were converted into percentile scores using a cumulative distribution function. Inferential statistics were computed using the t-test method developed by Crawford \& Howell, as described in \citet{Crawford_2009}. Two-sided p-values were computed based on the t-value and the n-1 degrees of freedom (n being the size of the control sample). All computations were implemented using in-house software written in Python 2.7. A heatmap visualizing the differences of JB’s z-scores in all major tasks was created using Seaborn (seaborn.pydata.org).
